\section*{Biographical sketches}
\label{sec:bios}

\paragraph*{Kai Xu} received his PhD in Computer Science at the National University of Defense Technology (NUDT).
He is now a faculty member at NUDT and a visiting associate professor at SIAT.
Between 2009 and 2010, he visited Simon Fraser University, supported by the Chinese government. His research interests include geometry processing and geometric modeling, especially methods that utilize large collections of 3D shapes.
He is currently serving on the editorial board of Computer and Graphics.
He has served on the program committees of SGP 2013--2015, PG 2013--2015 and GMP.

%\vspace{-.3cm}

\paragraph*{Vladimir G. Kim} is a Research Scientist at Adobe Research. His interests include geometry processing and analysis of shapes and collections of 3D models. He received his PhD in the Computer Science Department at Princeton University in 2013, and was a postdoctoral scholar at Stanford University 2013--2015. Vladimir is a recipient of the Siebel Scholarship and the NSERC Postgraduate Scholarship. He was also on the International Program Committee for SIGGRAPH 2015, SGP 2013--2015, and PG.

%\vspace{-.3cm}

\paragraph*{Qixing Huang} is a research assistant professor at TTI Chicago. He earned his PhD from the Department of Computer Science at Stanford University in 2012. He obtained both his MS and BS degrees in Computer Science from Tsinghua University in 2005 and 2002, respectively. His research interests include data-driven geometry processing and co-analysis of shapes and collections of 3D models using convex optimization techniques. He received a Best Paper Award from SGP 2013 and the Most Cited Paper Award from the journal of Computer-Aided Geometric Design in 2011 and 2012. He has served on the program committees of SGP, PG and GMP.

%\vspace{-.3cm}

\paragraph*{Evangelos Kalogerakis} is an assistant professor in computer science at the University of Massachusetts Amherst. His research deals with automated analysis and synthesis of 3D visual content, with particular emphasis on machine learning techniques that learn to perform these tasks by combining data, probabilistic models, and prior knowledge. He obtained his PhD from the University of Toronto in 2010 and BEng from the Technical University of Crete in 2005. He was a postdoctoral researcher at Stanford University from 2010 to 2012.  He served on the program committees of EG 2014 and 2015, SGP 2012, 2014 and 2015. His research is supported by NSF and NVidia.

%\vspace{-.3cm}

\section*{Acknowledgements}
We thank Zimo Li for proofreading this survey and the anonymous reviewers for helpful suggestions.
Kalogerakis gratefully acknowledges support from NSF (CHS-1422441).
Kai Xu is supported by NSFC (61572507, 61202333 and 61532003).
